---
title: 'nausea: Nowcasting Under Structural Breaks'
output:
  rmarkdown::html_vignette: default
  html_document:
    df_print: paged
  pdf_document: default
vignette: |
  %\VignetteIndexEntry{nausea} %\VignetteEngine{knitr::rmarkdown} \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r} 
library("nausea")
library("nowcasting")
```

## Getting data

`get_data` will download current vintage data from the NYFED and FREDMD databases up to a given date, defaulting to June 2021. These are subjected to stationarity transforms.

```{r}
data_vignette <- get_data()
```

By default, we include a version up to May 2021.
```{r}
data(panel)
```

## Factor modelling

`factor_model` will fit a factor model, selecting the factor number using information criterion #2 from Bai and Ng (2002).
An object of class `factor_model` is returned, which has `summary`, `plot` and `predict` methods.

```{r}
factor_model_out <- factor_model(panel$panel )
r_ <- factor_model_out$q.hat
factors <- factor_model_out$f.q#[,1:r_]
summary(factor_model_out)
```
```{r}
plot(factor_model_out)
```


## Change point analysis with `mosumvar`

Based on the `mosumvar` package and methodology, `mosumvar_factor` will perform change point analysis on the panel assuming a static factor model with low-dimensional vector autoregression (VAR) dynamics on the factors. The order `p` is recommended to be 1. The bandwidth `G` needs to be large enough for estimation purposes, but small enough to capture the quickly changing structure of the data, which is a difficult problem with monthly data. We recommend `G=18` or `24`. For similar reasons, we recommend the `Score` method. `nu` quantifies how close successive change points are allowed to be declared in the estimation phase; `nu=.1` is a suitably flexible choice. 

```{r}
mosumvar_out <- mosumvar_factor(panel$panel, p = 1, G= 24, nu = .1, method = "Score")
```

## Pooled forecasting

`pooled_forecast` will forecast the factor series using the most recent estimated change point, using either `robust`,`exp`, or `equal` weights on the models fit prior to the change. `predict` will predict for the panel.

```{r}
pooled_forecast_out <- pooled_forecast(factors, cp = mosumvar_out$cps, p =1, window_size = 50, weights = "robust", n.ahead = 6)
plot(pooled_forecast_out, main = "factor forecast")
```
```{r}
panel_forecast <- predict(factor_model_out, pooled_forecast_out)
```


## Nowcasting

Nowcasting of the GDP component can be performed with e.g. a linear regression model via `lm`, which permits use of the `predict` method. 

```{r}
lm_data <- data.frame(gdp = panel$gdp, factors)
gdp_lm <- lm(gdp ~ ., data = lm_data)
gdp_lm
plot.ts(gdp_lm$fitted.values, xlab = "Time (Quarter)", ylab = "smoothed change in gdp")
predict(gdp_lm, newdata = data.frame(gdp = NA, pooled_forecast_out))
```


